---
title: "Plotting Andro sensor app data using R"
output: html_notebook
---

### Objective/s 
Requirements CSV files created using [AndroSensor] Android App.  
1. Install required packages (rgdal, sp)  
2. Read all csv files from Andro Sensor Android app  
3. Create ESRI point shapefile from CSV  
 
### 1. Install Packages

```{r}
install.packages(c("rgdal", "sp"))
```

### 2. Read all CSV files created by Andro sensor Android app

- Set working directory  
```{r}
setwd("processAdroSensor")
```
- Folder with csv files
```{r}
csvFolder <- "drive-download-20160923T094238Z\\"
datFiles <- list.files(csvFolder)
```

- Read each CSV and add to dataframe
```{r}
all = data.frame()
for(i in c(1:length(datFiles))){
  #print(datFiles[i])
  f = read.csv(paste(csvFolder, datFiles[i], sep=""), header=T, sep=",")
  #names(f)
  colnames(f) <- c( "ACC.X","ACC.Y","ACC.Z","G.X","G.Y","G.Z","L.A.X", "L.A.Y","L.A.Z", "GYRO.X","GYRO.Y", "GYRO.Z","LIGHT","MAG.X","MAG.Y","MAG.Z","OR.Z", "OR.X", "OR.Y", "PROXIMITY", "S.LEVEL","Lat","Lon", "Alt", "Alt.g","Speed", "Accuracy", "ORIENT", "Sat.in.range", "T.since.start","date")
  all <- rbind(all, f)
}
head(all)
```

- print class of each column 

```{r}
lapply(1:length(all), function(x)class(all[,x]))
```

- Select required columns
```{r}
dataOut <- all[,c(22,23,24,25,27,29,31)]
names(dataOut)
dim(dataOut)
```
  
- Remove NA values  
```{r}
dataOut <- na.omit(dataOut)
dim(dataOut)
```

- Convert Date from factor to character to Date class
```{r}
dataOut$date <- as.Date(as.character(dataOut$date))
print(head(dataOut$date))
```

- Print class of each column

```{r}
lapply(1:dim(dataOut)[2], function(x)class(dataOut[,x]))
```

- (Optional) Write formatted data to CSV file 

```{r}
write.csv(dataOut, "allData3.csv", row.names = FALSE)
```

- Load required R packages
```{r}
library('sp')
library('rgdal')
```

- Verify class it should be data.frame

```{r}
class(dataOut); names(dataOut)
```

- Set lat lon columns from data.frame

```{r}
coordinates(dataOut)<-~Lon+Lat 
```

- check class and column names to know sp is associated

```{r}
class(dataOut)
names(dataOut)
summary(dataOut)
```

- Add projection to shapefile

```{r}
proj4string(dataOut) <- CRS("+init=epsg:4326")
```

### 3. Create ESRI point shapefile from CSV 
- (Optional) Write shapefile
```{r}
writeOGR(dataOut, ".", "fieldVisitLoc", driver = "ESRI Shapefile")
``` 

- Sample spplot 1
```{r}
spplot(dataOut, "Alt", do.log = FALSE,
       key.space=list(x=0.2,y=0.9,corner=c(0,1)),
       scales=list(draw=T))
```

- Sample spplot 2

```{r}
spplot(dataOut, "Alt.g", do.log = FALSE,
       key.space=list(x=0.2,y=0.9,corner=c(0,1)),
       scales=list(draw=T))
```

### References
- [R  SP Package]
- [R  rgdal Package]
- [AndroSensor]

[R  SP Package]:<https://cran.r-project.org/web/packages/sp/index.html>
[R  rgdal Package]:<https://cran.r-project.org/web/packages/rgdal/index.html>
[AndroSensor]:<https://play.google.com/store/apps/details?id=com.fivasim.androsensor&hl=en>

### Author details

- Author: Suryakant Sawant, PhD
- Date: 24 Sept. 2016
